### 2018-12-18 README.Rmd
Initial commit: Setup repository

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# fortebaseline

This repository contains the source code and raw manuscript files for the Hector permafrost emissions paper.

## Installation

You can install the development version of `hector_permafrost_emit` from GitHub with:

```r
# install.packages("devtools")
devtools::install_github("ashiklom/hector_permafrost_emit")
```