### 2020-03-17 README.Rmd
Update README with renv instructions

---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# Hector permafrost emissions analysis

This repository contains the source code and raw manuscript files for the Hector permafrost emissions paper.
The paper source code is located in `analysis/paper/paper.Rmd`.

## `renv`

This repository uses `renv` to manage R package dependencies in a reproducible way.
To download and install all packages in the correct versions for this analysis, launch R from the project root directory and run `renv::restore()`.